\LARGE \bf \pyRobots, a toolkit for robot executive control

### 2014-08-22 paper.tex
Work on events, resource and poses sections

\usepackage[protrusion=true,expansion=true]{microtype}
\usepackage{cite}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{url}

@@

\begin{itemize}
    \item Turns any Python function into a background action with the decorator
        \python|@action|.
    \item Robot actions are non-blocking by default: they are instanciated as
        futures (lightweight threads),
    \item Actions can be cancelled at any time via signals (the
        \python|ActionCancelled| signal is raised).
    \item Lock specific resources with a simple \python|@lock(...)| in front of the
        actions. When starting, actions will wait for resources to be
        available if needed.
    \item Supports compound resources (like \python|WHEELS == LEFTWHEEL + RIGHTWHEEL|)
    \item Create event with \python|robot.every(<condition>).do(<action>)|
    \item Poses are managed explicitly and can easily be
        transformed from one reference frame to another one
        (integrates with ROS TF when available).
    \item Extensive logging support to debug and replay
        experiments.
\end{itemize}

@@

\pyRobots core concepts are simple: on one hand, the user creates a
\textbf{robot} as an instance of a \python|GenericRobot|, that encapsulates the
different low-level \textbf{controllers} (proxies to actuators, typically
provided by a middleware like ROS) and the \textbf{state} of the robot
(typically a set proxies to the sensors, or a connector to a database, etc.).

@@

\emph{Actions} in \pyRobots are always asynchronous. They are implemented as
\emph{futures} (also known as \emph{promises}) and therefore executed in
independent threads. To the developer, however, they appear as regular Python
function, simply annotated with the \python|@action| decorator, and are invoked
as any other function.  Since they are asynchronous, these functions may
perfectly block or even never return (typically useful to implement lasting
background behaviours).

Actions can also call each-other, allowing natural decomposition of complex
tasks into sub-actions and eventually atomic actions.

\paragraph{Signaling} Running behaviours often need to be paused of cancelled.
This poses a particular challenge in robotics, because before suspending an
action, one often want to bring back the robot in some form of rest state (you
would not want to suddenly cancel a \python|walk| action in the middle of a step
without first bringing back the foot to the ground).

\pyRobots addresses this issue by the mean of signals, and extends standard
Python thread into \emph{signaling threads}. Listing~\ref{lst:signals}
illustrates this mechanism.

\begin{listing}[H]
\begin{pythoncode}
    @action
    def my_action(robot):
      try:
        robot.do_something_dangerous()
      except ActionCancelled:
        robot.go_back_to_rest_pose()

@@

@@

@@

\subsection{Uniform management of rigid transformations and frames}

@@

\section{Implementation and Example}

%\begin{listing}[ht]

@@

        # create (and set) one element in the robot's state.
        # Here a bumper.
        self.state.my_bumper = False

@@

      robot.whenever("my_bumper", value = True)
                                    .do(move_forward)

      try:
        while True:
          time.sleep(0.5)
      except KeyboardInterrupt:
        pass
\end{pythoncode}
%\caption{Example from external file}
%\label{listing:3}
%\end{listing}

>>>>

\usepackage[protrusion=true,expansion=true]{microtype}
\usepackage{cite}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{url}
\usepackage{amsmath}

@@

\begin{itemize}
    \item Lock specific resources with a simple \python|@lock(...)| in front of the
        actions. When starting, actions will wait for resources to be
        available if needed.
    \item Supports compound resources (like \python|WHEELS == LEFTWHEEL + RIGHTWHEEL|)
    \item Poses are managed explicitly and can easily be
        transformed from one reference frame to another one
        (integrates with ROS TF when available).
    \item Extensive logging support to debug and replay
        experiments.
\end{itemize}

@@

\pyRobots core concepts are simple: on one hand, the user creates a
\textbf{robot} as an instance of a \python|GenericRobot|, that encapsulates the
different low-level \textbf{controllers} (proxies to actuators, typically
provided by a middleware like ROS) and the \textbf{state} of the robot
(typically a set proxies to the sensors, or a connector to a database, etc.,
accessed through a dictionary-like object).

@@

\emph{Actions} in \pyRobots are always asynchronous. They are implemented as
\emph{futures} (also known as \emph{promises}) and therefore executed in
independent threads. To the developer, however, they appear as regular Python
function, simply annotated with the \python|@action| decorator, and are invoked
like any other function.  Since they are asynchronous, these functions may
perfectly block or even never return (typically useful to implement lasting
background behaviours).

Actions can also call each other, allowing natural decomposition of complex
tasks into sub-actions and atomic actions.

\paragraph{Signaling} Behaviours often need to be paused of cancelled.  This
raises an issue that needs to be addressed in robotics: before suspending an
action, one often wants to first bring back the robot into some form of rest
state (\eg one would not want to suddenly cancel a \python|walk| action in the
middle of a step without first bringing back the foot to the ground).

\pyRobots addresses this issue through signals, and extends standard
Python thread into \emph{signaling threads}, relying on Python's usual syntax
for exception handling. Listing~\ref{lst:signals}
illustrates this mechanism. The action is instantiated and started at line 8,
and cancelled at line 10. This send a \python|ActionCancelled| signal to the
action that the user can handle using an \python|except| statement.

\begin{listing}[H]
\begin{pythoncode}
    @action
    def my_action(robot):
      try:
        robot.do_something_long()
      except ActionCancelled:
        robot.go_back_to_rest_pose()

@@

Pausing actions is more complex (the action internal state must be saved, a mean
to resume the action where it was interrupted must be provided, the user must be
given ways to code transition behaviours when pausing and resuming) and is not
currently supported by \pyRobots.

@@

Sharing the same syntax with URBI, events in \pyRobots can by monitored through
the methods \python|on| (one shot event) and \python|whenever| (continuous event
monitoring). Event conditions can be either a predicate (that must take as
single argument the robot instance) or a pair of $\{\text{state member},
\text{condition}\}$ with condition in $\{\text{value}=x, \text{become}=x,
\text{below}=x, \text{above}=x, \text{decrease}=x, \text{increase}=x\}$. An
example of event usage can be seen at line 67 of listing~\ref{mwe}.

When the event condition is verified, the callback (usually, an action) is
invoked from the event monitor thread.

@@

Usually, physical robot resources (mainly actuators, but possibly sensors as
well) require exclusive use (it is seldom desirable that two different actions
control the wheels of a robot at the same time).

To leave full control regarding the required resource modeling granularity, the
programmer define them himself, either as stand-alone resources or as compound
resources (listing~\ref{lst:resources}).

\begin{listing}[H]
\begin{pythoncode}
    L_ARM = Resource()
    R_ARM = Resource()
    ARMS = CompoundResource(L_ARM, R_ARM)

    @action
    @lock(ARMS)
    def lift_box(robot):
        #...

    @action
    @lock(L_ARM)
    def wave_hand(robot):
        #...

    @action
    @lock(L_ARM, wait=False)
    def scratch_head(robot):
        #...

    robot.lift_box()
    robot.wave_hand() # waits until lift_box is over
    robot.scratch_head() # skipped

\end{pythoncode}
\caption{Resource usage is defined at the action-level, through annotations.}
\label{lst:resources}
\end{listing}

The user then annotates the actions with the (possibly multiple) resources they
require (lines 6, 11 and 16 of listing~\ref{lst:resources}). When executed, the
action acquires a lock on the resource (and all dependent resources in the case
of a compound resource) which is released upon action completion. In the
meantime, other actions requiring the same resource will either wait until the
resource is release (line 21) or be skipped (line 22) if the \python|@lock|
annotation specifies not to wait (line 16).

\subsection{Uniform management of poses and rigid transformations}

Managing poses, transformations, frames in real-world scenarios is not an easy
task, mainly because it seems that every single library, middleware or robotic
platform rely on its own conventions (rotation representations, in particular,
are a constant source of frustration, having to jump all the time between
quaternions, one of the 23 Euler conventions, rotation matrices, Rodrigues,
yaw-pitch-roll notation...). Within the ROS ecosystem, the Transform Library
(tf,~\cite{foote2013tf}) has played a major role to make it convenient to work
with frames, and poses can always be expressed and shared in the most convenient
local frame, tf taking care of required intermediate transformation.

\pyRobots offers a similar approach \emph{within} the control script, adding a
pinch of syntactic sugar to make poses and frame management transparent for the
high-level action controller.

As illustrated in listing~\ref{lst:poses}, poses can be flexibly defined by a
simple frame name, by a triple $\{x, y, z\}$ (default frame {\tt map} is then
assumed), by a 4-tuple $\{x, y, z, frame\}$, by a 6-tuple $\{x, y, z, rx, ry,
rz\}$ (using $sxyz$ Euler convention), by a 7-tuple $\{x, y, z, rx, ry, rz,
frame\}$, by a 7-tuple $\{x, y, z, qx, qy, qz, qw\}$, by a 8-tuple $\{x, y, z,
qx, qy, qz, qw, frame\}$ or by a dictionary that defines any of these 8 values.
Internally, \pyRobots normalizes and stores the poses in their full 8-tuple
expansion.

\begin{listing}[H]
    \begin{pythoncode*}{linenos=false, xleftmargin=0em}

    # {'x':0., 'y':0., 'z':0., 
    #  'qx':0., 'qy':0., 'qz':0., 'qw':1., 
    #  'frame': 'handL'}
    robot.poses["handL"]

    # {'x':1., 'y':0., 'z':0., 
    #  'qx':0., 'qy':0., 'qz':0., 'qw':1., 
    #  'frame': 'map'}
    robot.poses[(1.0, 0.0, 0.0)]

    # {'x':0., 'y':0., 'z':1., 
    #  'qx':0., 'qy':0., 'qz':0.0998, 'qw':0.0995, 
    #  'frame': 'handL'}
    robot.poses[{'z':1.0, 'rz':0.2, 'frame':'handL'}]

\end{pythoncode*}
\caption{Examples of poses normalizing.}
\label{lst:poses}
\end{listing}

Using helper methods like \python|robot.poses.inframe(<pose>, <other frame>)| or
\python|robot.poses.pantilt(<pose>, <ref frame>)|, the programmer can then
transform a given pose into a different frame or different representation.

These transformations are directly computed, but \pyRobots also takes advantage
of middleware-specific infrastructure (when available) to provide known frames.
ROS TF, Aldebaran's {\tt NAOqi} frame conventions, etc. are used when available
through specializations of the \pyRobots \python|PoseManager| class.

@@

\section{Detailed Example}

\begin{listing}[h!]

@@

        # create (and set) one element in the robot's state.
        # Here a sonar.
        self.state["sonar"] = 10.

@@

      robot.whenever("sonar", below = 0.3).do(move_forward)

      try:
        while True:
          time.sleep(0.5)
      except KeyboardInterrupt:
        pass
\end{pythoncode}
\caption{Full example of a \pyRobots application.}
\label{mwe}
\end{listing}
